The Practical Application of MST

### 2021-02-16 article.Rmd
Added simulation of single candidate with mstR

---
title: "The Practical Application of MST"
description: |
  A practical guide to design and development of computerised multistage tests.
author:
  - name: James Page
    affiliation: SP Educational Consulting
    affiliation_url: https://www.speduconsulting.co.uk/
date: "`r Sys.Date()`"
bibliography: references.bib
output: distill::distill_article
---

@@

Computerised multistage tests (MSTs) have grown in popularity in recent years due to their advantages of flexibility and simplicity over linear and computerised adaptive tests (CATs). A growing body of academic literature is available for researchers, but the practicalities of using MST for design and delivery of real-world tests is still difficult to find. I hope this article provides a useful guide to the practical work of test creation and delivery.

@@

I'll assume some familiarity with Item Response Theory (IRT) - and the Rasch model. Also some competence with R programming would be useful to understand the examples given. For background information and primers on these subjects I provide references.

@@

This S-shaped function can be seen below for three test items with different difficulties.

@@

```{r module difficulties}
library(mstR)

@@

```{r module items}
modules <- matrix(0, 68, 7)
modules[1:12, 1] <- 1
modules[13:20, 2] <- 1
modules[21:28, 3] <- 1
modules[29:38, 4] <- 1
modules[39:48, 5] <- 1
modules[49:58, 6] <- 1
modules[59:68, 7] <- 1
```

@@

```{r module routing}
routing <- matrix(0, 7, 7)
routing[1, 2:3] <- 1
routing[2, 4:6] <- 1
routing[3, 5:7] <- 1

print(routing)
```

@@

>>>>

---
title: "The Practical Application of MST"
description: |
  A practical guide to the design and development of computerised multistage tests.
author:
  - name: James Page
    affiliation: SP Educational Consulting
    affiliation_url: https://www.speduconsulting.co.uk/
date: "`r Sys.Date()`"
bibliography: references.bib
output: distill::distill_article
---

@@

Computerised multistage tests (MSTs) have grown in popularity in recent years due to their advantages of flexibility and simplicity over linear and computerised adaptive tests (CATs). A growing body of academic literature is available for researchers, however the practicalities of using MST for the design and delivery of real-world tests is still difficult to access I hope this article provides a useful practical guide to test design and delivery.

@@

I'll assume some familiarity with Item Response Theory (IRT) - and the Rasch model. Some knowledge of programming with R would be useful to understand the examples given. For background information and primers on these subjects I provide references.

@@

The sigmoid function can be seen below for three test items with different difficulties.

@@

```{r module difficulties, echo=TRUE}
library(mstR)

@@

```{r module items, echo=TRUE}
modules <- matrix(0, 68, 7)
modules[1:12, 1] <- 1
modules[13:20, 2] <- 1
modules[21:28, 3] <- 1
modules[29:38, 4] <- 1
modules[39:48, 5] <- 1
modules[49:58, 6] <- 1
modules[59:68, 7] <- 1
```

@@

```{r module routing, echo=TRUE}
routing <- matrix(0, 7, 7)
routing[1, 2:3] <- 1
routing[2, 4:6] <- 1
routing[3, 5:7] <- 1

cutoff <- matrix(NA, 7, 2)
cutoff[2,] <- c(-Inf, 0.0)
cutoff[3,] <- c(0.0, Inf)
cutoff[4,] <- c(-Inf, -1.0)
cutoff[5,] <- c(-1.0, 0.0)
cutoff[6,] <- c(0.0, 1.0)
cutoff[7,] <- c(1.0, Inf)

print(cutoff)
```

@@

Having created a simulated item bank for the design, it is possible to generate simulated item responses by selecting a $\theta$ for a simulated candidate:

```{r simulated test response, echo=TRUE}
theta <- -0.79

dfSampleResponse <- genPattern(theta, it.MST)

print(dfSampleResponse)
```

Here we use a $\theta$ of `r theta`. Note that this doesn't really simulate a true MST test, as it ignores the routing rules and simply returns a response for all items in the simulated item bank. We can however use the `randomMST()` function to perform a more realistic simulation. The `randomMST()` function will generate a response pattern for a given simulated item bank, an MST structure, and a true ability level ($\theta$). We'll use maximum likelihood estimation (`"ML"`) for ability estimation, and the maximum Fisher information criterion (`"MFI"`) for module selection.

```{r simulated MST test response, echo=TRUE}
start <- list(theta = theta)
test <- list(method = "ML", moduleSelect = "MFI")
final <- list(method = "ML")

dfMSTResponse <- randomMST(trueTheta = theta, 
                           itemBank = it.MST, modules = modules, transMatrix = routing,
                           start = start, test = test, final = final)

(dfMSTResponse)
```

The Module Information Function is used to determine the 'best' (ie. the most informative) module to route to in the next stage. In this example the selected modules are `r dfMSTResponse$selected.modules`. A plot can be made as follows, showing the highlighted route through the multi-stage test:

```{r show simulated test, echo=TRUE}
plot(dfMSTResponse)
```
